---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Effect of Mass Transport Systems and Economic output.

```{r}
library(stargazer)
library(tidyverse)
library(skimr)
library(ggplot2)
library(imputeTS)
library(did)
library(stargazer)
library(recipes)
library(did2s)
library(fixest)
```

```{r}
# load Santiago city data file 
stgo <- read_csv("/Users/hendrixperalta/Downloads/Projects/Will-A-Subway-Line-Reduce-Transit-related-Deaths/stgo.csv") 
```

```{r}
# load Santo Domingo city data file 
sd <- read_csv("/Users/hendrixperalta/Downloads/Projects/Will-A-Subway-Line-Reduce-Transit-related-Deaths/sd.csv") 

```

```{r}
# 2013 - 2020 period is choosen due to data availability 

sd_13_20 <- sd %>% filter(year >= 2013 & year <= 2020) 
```

```{r}
skim(sd_13_20) %>% as.tibble
```

```{r}
# Transform the variable year to numeric 
# Create a post variable indicating the treatment period 
# Create a post_factor variable that labels the periods of pre and post treatment
# Only the variables year, homicide, medical_checks, pop, ntl, pm25, transit_death, post and post_factor are selected

sd_fix <- sd_13_20 %>% 
  mutate(
    year = as.integer(year),
    post = ifelse(year >= 2018, 1, 0), 
    post_factor = factor(post , labels = c("Pre-Treatment", "Post-Treatment"))
  ) %>% 
  select(year, homicide, medical_checks, pop, ntl, pm25, transit_death, post, post_factor)
```

```{r}
# No missing values in the data set

skim(sd_fix) %>% as.tibble
```
### Linear Regression 

```{r}
# Create a box plot for each group of 'post'

ggplot(sd_fix, aes(x = post_factor, y = ntl)) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Transit-related Deaths",
       title = "Distribution of Transit-related Deaths by Post-Treatment")
```

We can observe that nightime lights present different distributions before and after the treatment. On that regard, the mean and median of nighttime Lights is higher in the post treatment group hinting a correlation between the introduction of mastransportation systems to the Santo Domingo city and economic output. 

```{r}
# Definition of the regression models

model_1 <- lm(ntl ~ post, data = sd_fix)
model_2 <- lm(ntl ~ post + medical_checks , data = sd_fix)
model_3 <- lm(ntl ~ post + medical_checks + transit_death, data = sd_fix)
model_4 <- lm(ntl ~ post + medical_checks + transit_death + homicide , data = sd_fix)
#model_5 <- lm(transit_death ~ post + medical_full + ntl_root + homicide_root + pm25, data = sd_fix)
#model_6 <- lm(transit_death ~ post + medical_full + ntl_root + homicide_root + pm25 + pop, data = sd_fix)


#ntl_root, homicide_root, medical_full, pop, pm25, post, transit_death
```

```{r}
residual_1 <- resid(model_1)
residual_2 <- resid(model_2)
residual_3 <- resid(model_3)
residual_4 <- resid(model_4)
#residual_5 <- resid(model_5)
#residual_6 <- resid(model_6)
```

```{r}
# Creates a plot to check for patterns in the regression residuals

ggplot(data = data.frame(residual_4)) +
  geom_point(aes(x = seq_along(residual_4), y = residual_4), pch = 16, col = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Observation", y = "Residuals", title = "Residual Plot for NTL Model")
```
there are no visible patterns on the residuals of the regression indicating that there is no heteroschedasticity on our dataset. 


```{r}
# Regression table - includes the 4 regression models

variable_names <- c(
  "post" = "Post Subway",
  "medical_checks" = "Access to Healthcare",
  "transit_death" = "Transport Related Deaths",
  "homicide" = "Social Insecurity"
)

stargazer(model_1, model_2, model_3, model_4,
          type = "text", covariate.labels = variable_names,
          notes = c("GDP = Masked Mean Nighttime Lights", 
                    "Pollution = Count of Particule Matter PM 2.5", 
                    "Access to Healthcare = Count of Medical Checks in Public Facilities", 
                    "Social Insecurity = Count of homicides cases"))
```

In the table we can observe that the post dummy variable indicating the introduction of the mass transportation system have big and statistical significance in all 4 models. 


### Difference in Differences 

```{r}
# 2013 - 2020 period is choosen due to data availability 

stgo_13_20 <- stgo %>% filter(year >= 2013 & year <= 2020)
```

```{r}
stgo_fix <- stgo_13_20 %>% 
  
# Transform the variable year to numeric 
# Create a "post" variable indicating the treatment period 
# Create a "post_factor" variable that labels the periods of pre and post treatment
# Only the variables year, homicide, medical_checks, pop, ntl, pm25, transit_death, post and post_factor are selected
  
  mutate(
    year = as.integer(year),
    post = ifelse(year >= 2018, 1, 0), 
    post_factor = factor(post , labels = c("Pre-Treatment", "Post-Treatment"))
  ) %>% 
  select(year, homicide, medical_checks, pop, ntl, pm25, transit_death, post, post_factor)
```

```{r}

# Create variable "treat" to indicate if it is the treatment or control group (used with did library)
# Craate variable "id" to indicate if it is the treatment or control group (used with fixest and did2s libraries)
# Scale variables bt population to make them comparable

sd_did <- sd_fix %>% mutate(
    treat = TRUE,
    id = "santo Domingo",
    year = as.integer(year),
    death_pop = transit_death/(pop/1000),
    homicide_pop = homicide/(pop/1000),
    medical_pop = medical_checks/(pop/1000)
)

stgo_did <- stgo_fix %>% mutate(
    treat = FALSE,
    id = "santiago",
    year = as.integer(year),
    death_pop = transit_death/(pop/1000),
    homicide_pop = homicide/(pop/1000),
    medical_pop = medical_checks/(pop/1000)
)
```

```{r}


ggplot()+ 
geom_line(data = sd_did, aes(x = year, y = ntl, color = "Santo Domingo", ))+ 
geom_line(data = stgo_did, aes(x = year, y = ntl, color = "Santiago")) +
geom_vline(xintercept = 2018, linetype = "dashed", color =  "black") + 
xlab("Year") +
  ylab("Nighttime Lights") +
  ggtitle("NTL Increased After Introducing the cable car and Subway Line expansion") +
  theme_minimal()
```

```{r}
combined_data <- bind_rows(sd_did, stgo_did)
```

```{r}
combined_data
```

```{r}
skim(combined_data)
```

```{r}
model_did = lm(ntl ~ treat + post + treat * post, data = combined_data)
model_did_2 = lm(ntl ~ treat + post + medical_pop + death_pop + homicide_pop + treat * post, data = combined_data)
```

```{r}

stargazer(model_did, model_did_2, type = "text")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or
by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output
will be saved alongside it (click the *Preview* button or press
*Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the
editor. Consequently, unlike *Knit*, *Preview* does not run any R code
chunks. Instead, the output of the chunk when it was last run in the
editor is displayed.

```{r}
sd_did2 <- sd_fix %>% mutate(
    treat = TRUE,
    id = 1,
    first_treat = 2018,
    rel_time = year - 2018,
    year = as.integer(year),
    death_pop = transit_death/(pop/1000),
    homicide_pop = homicide/(pop/1000),
    medical_pop = medical_checks/(pop/1000)
)

stgo_did2 <- stgo_fix %>% mutate(
    treat = FALSE,
    first_treat = 0,
    rel_time = year - 2018,
    id = 0,
    year = as.integer(year),
    death_pop = transit_death/(pop/1000),
    homicide_pop = homicide/(pop/1000),
    medical_pop = medical_checks/(pop/1000)
)
```

```{r}
data_did2 <- bind_rows(sd_did2, stgo_did2) 
```

```{r}
data_did2 
```

```{r}
# did2s - Gardner 2021 - paper Two-stage differences in differences
es_mod = did2s(
  data         = data_did2,
  yname        = "ntl", 
  first_stage  =  ~ medical_pop + death_pop + homicide_pop | id + year, # 0 b/c we have no controls in this dataset
  second_stage = ~ i(rel_time, ref = -c(1, Inf)),      # binary treatment dummy (not an event-study) 
  treatment    = "post",
  cluster_var  = "id",
  )
```

```{r}
es_mod
```

```{r}
es_mod |>
  fixest::iplot(
    main     = "did2s",
    xlab     = "Time to treatment",
    drop     = "[[:digit:]]{2}",    # Drop any leads/lags greater than |9|
    ref.line = 1
  )
```

```{r}
skim(data_did2)

```

```{r}
# did - Callaway and Sant’Anna 2021 -  paper Difference-in-Differences with multiple time periods
cs21 = att_gt(
    yname         = "ntl",
    tname         = "year",
    idname        = "id",
    gname         = "first_treat",
  # xformla       = NULL,            # No additional controls in this dataset 
    control_group = "nevertreated", # Too few groups for "nevertreated" default
    data          = data_did2
    )
```

```{r}
atts <- att_gt(yname = "ntl", # LHS variable
               tname = "year", # time variable
               idname = "id", # id variable
               gname = "first_treat", # first treatment period variable
               data = data_did2, # data
               xformla = NULL, # no covariates
               #xformla = ~ l_police, # with covariates
               est_method = "dr", # "dr" is doubly robust. "ipw" is inverse probability weighting. "reg" is regression
               control_group = "nevertreated", # set the comparison group which is either "nevertreated" or "notyettreated" 
               bstrap = TRUE, # if TRUE compute bootstrapped SE
               biters = 1000, # number of bootstrap iterations
               print_details = FALSE, # if TRUE, print detailed results
               clustervars = "id", # cluster level
               panel = TRUE) # whether the data is panel or repeated cross-sectional
```

```{r}
# fixest - Sun and Abraham 2020 - paper Estimating Dynamic Treatment Effects in Event Studies with Heterogeneous Treatment Effects
sa20 = feols(
    ntl ~ sunab(first_treat, rel_time) | id + year, 
    data = data_did2, vcov = ~id
    )
sa20
```

\`\`\`{r} sa20 \|\> iplot( main = "fixest::sunab", xlab = "Time to
treatment", drop = "[[:digit:]]{2}", \# Limit lead and lag periods to
-9:9 ref.line = 1 )
